Using reference models in feature selection problems \vspace{.1in}

### 2019-07-24 paper/ref_approach.tex
edits

\begin{document}
\maketitle
\begin{abstract}

Feature selection, or more generally, model reduction is an important aspect of statistical workflows aiming to provide insides from data. While a vast number of feature selection procedures are available, it often remains unclear what are the underlying principles explaining why certain methods are working better than others. We investigate and discuss one of these principles in detail, which we identify with the use of a reference model. We demonstrate how such a model acts as a noise-filter on the target variable by approximating its data generation mechanism.
To this end, we devise a simple reference model approach that can be used on top of any feature selection method. In several numerical experiments, we show how this translates into better and more stable feature selection independently of the specific selection procedure. 
These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.

@@

In statistical applications, one of the main steps in the modelling
workflow is covariate or feature selection, which is a special case of
model reduction. Variable selection may have multiple goals. 
First, if the covariates themselves are of interest,
we can use variable selection as an inferential tool to find the entire 
collection of variables which have some predictive information about the target variable.
Second, as simpler models come with the advantages of reduced
measurement costs and improved interpretability, we may be interested in
finding the minimal subset of covariates which
provides good predictive performance (or good balance between simplicity and
predictive performance). 
Whenever the predictive capability is guiding the selection, the true data generation
mechanism of future data can be approximated either by using the observed data directly or
alternatively by using predictions from a \emph{reference model}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward 
regression \citep{venables2013modern,harrell2015regression}, the observed empirical data distribution is utilised as a
proxy of future data usually in combination with cross-validation
or information criteria to provide estimates for out-of-sample
predictive performance.
In contrast, reference model based methods approximate the future data generation
mechanism using the predictive distribution of a reference model,
which can be, for example, a full-encompassing model including all
covariates. The reference
model approach has been used for long time in Bayesian statistics
since the seminal work of \citet{paper:reference_lindley}. For more
historical references, see the review by \citet{vehtari2012survey}, and for
most recent methodological development see 
\citet{paper:projpred}.
Reference models have been also used in non-Bayesian contexts,
where \cite{harrell2015regression} describes them as full models that
can be thought of as a ``gold standard'' (for a given application).
For example,
\cite{faraggi2001understanding} deal with the necessity of identifying
interpretable risk groups in the context of survival data using neural
networks, which typically perform very well in terms of prediction,
but whose covariates are difficult to be understood in terms of
relevance. 
\cite{paul2008preconditioning}, using the term preconditioning,
explore approximating models fitting Lasso or stepwise regression
against consistent estimates $\hat{y}$ of a reference model instead of
the observed responses $y$.

In this paper, we examine the benefits of using reference
models in detail. We illustrate how a properly designed reference model is able 
to filter parts of the noise present in the data, and hence provides an improved and more
stable selection process. This holds even if the reference model does not actually resemble the
true data generating process. Moreover, our analyses indicate that the substantial reduction of
variance attributable to noise is usually
more important than small potential bias due model misspecification.
We argue that, regardless of how the reference model is set up and used in the
inference procedure, it can be always seen as acting as a filter on the observed 
data. Following this idea, we devise a simple reference model approach that can
be applied on top of any variable selection procedure.

In our simulations and analysis of real data, we apply several state-of-the-art methods for variable selection and compare,
for each of these methods, the results obtained by using a data based and a
reference model based approach. When a full 
variable selection is carried out, we report and compare quantities such as 
the false discovery rate \citep{paper:fdr_BH}, sensitivity and stability of the selection \citep{paper:stability}. For fully Bayesian methods, we additionally report results in terms of shrinkage and accuracy of the posterior estimates.
We show the benefits
of a reference model approach, regardless of what specific procedure is applied, using 
the normal means estimation problem as a benchmark. 
Variable selection is typically most challenging in 
\textit{small n large p} scenarios, that is, scenarios with small number of
observations but a lot of unknown model parameters. 
These occur frequently, for example, 
in microarray data analysis in medicine or biology \citep{dudoit2003multiple,liao2007logistic,paper:efron}.
In general, the less data we have and the more complex 
the estimated models, the higher the benefit of using a reference
models as the basis for variable selection.
Our results indicate that the core reason why the reference model based
methods perform well is the reference model itself, rather than the
specific way of using it. We do recognise that all these methods
diversify how the subset of variables is actually chosen and which
properties it has, thus it remains fundamental to investigate and
develop such different procedures. We want to note that, although we focus on
Bayesian approach in this paper, the benefits of the reference model 
approach can also be obtained in non-Bayesian settings.

@@

The main assumption under any reference model approach is that we
operate in an $\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}. That is, we assume to be able to construct a model whose predictions reflect the best available description of the uncertainty of future data. In practice, this will often be the full-encompassing model
(including all models we consider having non-zero prior
probability as submodels) which has passed model checking and criticism \citep[see,
e.g.][]{gelman2013bayesian}.

A good predictive model is able to filter parts of the noise present in the data. The noise is the main source of the instability and tends to obscure the relevance of the covariates in relation to the target variable of interest. Indeed, let us consider the following explanatory example taken from \cite{paper:projpred}. The data generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim N(0,1) && \nonumber \\ 
     Y|&f\sim N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} N(\sqrt{\rho}f,1-\rho) \quad &&j=1,..,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} N(0,1) &&j=k+1,..,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy observation. The first $k$ covariates are strongly related to the target variable $Y$ and correlated among themselves. Precisely, $\rho$ is the correlation among any pair of the first $k$ covariates, whereas $\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between any relevant covariate and, respectively, $f$ and $Y$. If we had an infinite amount of observations, the sample correlation would be equal to the true correlation between $X_j$ and $Y$. However, even in this ideal asymptotic regime, this  correlation would still remain a biased indicator of the true relevance of each covariate (represented by the correlation between $X_j$ and $f$) due to the intrinsic noisy nature of $Y$.

@@

% In Figure \ref{fig:correlation} the reference model filters the noise.
% This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
% Above the correlation was used to estimate the relevance of a covariate.  
As an another example, let us consider a candidate submodel in a feature selection process and let us refer to the parameter distribution of the submodel by $\pi$ and to the induced predictive distribution by $q_{\pi}(\tilde{y})$. We would like to choose $\pi$ so that $q_{\pi}(\tilde{y})$ maximises some predictive performance utility, for example, the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y},
\end{equation}
where $p_{t}(\tilde{y})$ denotes the (usually unknown) true generating mechanism of future data $\tilde{y}$. If we refer to the posterior predictive distribution of a reference model with $p(\tilde{y}|D)$, where $D$ stands for the data on which we conditioned on, we can approximate \eqref{eq:elpd} using $p(\tilde{y}|D)$ instead of the true data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the KL divergence between the reference model's and the submodel's predictive distributions:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred to as the projection of the predictive distribution, which is the general idea behind the projection predictive approach \cite[see][]{paper:projpred}. Again the reference model is acting as a filter, or substitute, on data.

@@

Relying on the idea of the reference model as a noise-filter, a generic reference model approach, which can be used on top of any feature selection procedure, consists of simply substituting the target observations with corresponding point predictions of the reference model, for instance, the posterior predictive means. Afterwards, any method for feature selection can be used. Note that we do not argue that such a reference model approach should necessarily be used as is for feature selection in real applications. There are several more sophisticated reference model approaches available in the literature \cite[see, e.g.][]{paper:projpred,paananen2017variable,paul2008preconditioning}, which are likely to yield even better results. Rather, the reason for focussing on this simple version is to enable a fair comparison between a reference model and a purely data based approach when otherwise using the same selection procedures. Other kinds of comparisons, as the one presented in Section \ref{projection}, work well as motivational examples but fail in terms of generality since the presence or absence of a reference model is not the only aspect differentiating the two procedures. We will come back to the more general comparison in Section \ref{comparison}.

@@

As a motivational example to the implications of using a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples to the submodels \citep{paper:original_proj}. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Suppose we have observed $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates for which we want to obtain a minimally relevant subset. Than, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of $S$ posterior samples of the reference model's parameters.
\item Rank the covariates according to their relevance using some heuristics and consider as candidate submodels only those which preserve this order,  starting from including only the highest ranked covariate (the submodels are then naturally identified by their model size). This is not strictly necessary but reduces the number of submodels to consider in the following steps.
\item For each submodel, project the reference model's samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL} \left( p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\;||\;q(\tilde{y}_{i}|\boldsymbol{\theta}^{s}) \right) \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the predictive distribution of the reference model with parameters fixed at the posterior sample $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariate values related to the statistical unit (identified by the subscript $i$), whereas $q(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ is the predictive distribution of the submodel. The projected samples $\boldsymbol{\theta}_{\perp}^{s}$ then correspond to the submodel's parameter distribution.
\item For each submodel (size), test the predictive performance for instance via cross-validation.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model's predictive utility score.
\end{enumerate}

@@

We show the benefits of the reference model in the form of the projection approach (projpred) with respect to the stepwise backward regression (steplm) using the body fat data \citep{johnson1996fitting}. The analysis is based on Vehtari's \texttt{R}-notebook available at \url{https://avehtari.github.io/modelselection/bodyfat.html}, which is inspired by \cite{paper:bodyfat}. The target variable of interest is the amount of body fat, which is measured with a complex and expensive procedure. Additionally, we have information about 13 covariates which are anthropometric measurements (e.g., height and weight), some of which are highly correlated among themselves and thus provide some challenges in the model selection. In total, we have 251 observations. The goal is to find the simplest model which is able to predict the amount of fat sufficiently well. We use the same data again later on in Section \ref{real-world-data}.

\cite{paper:bodyfat} report results using steplm with a significance level of 0.157 with AIC selection \citep{akaike1974new}, fixing abdomen and height to be always included in the model. Here, we implement steplm via AIC without fixing any variable. For the selection via projpred, the reference model includes all the covariates, we regularise model complexity with a regularised horseshoe prior on the covariate coefficients \citep{paper:rhs}, submodels are explored using forward search, and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. The experiments are implemented in \texttt{R} \citep{Rcore2018}. The steplm approach is carried out combining the \texttt{step} and \texttt{lm} functions, whereas the selection via projpred is done with the \texttt{projpred} package. In order to automatise the latter, we select the submodel as the one with the smallest size which has an elpd score higher (i.e., better) than the reference model score with probability of at least 5\%. The inclusion frequencies of each feature in the final model after 100 bootstrap samples are depicted in Figure \ref{fig:inclusion_frequencies}. The projpred approach has only two variables, `abdomen' and `weight', with inclusion frequencies above 50\% (`abdomen' is the only included always), the third most included is `wrist' at 44\%, and the fourth is `height' at 35\%, while steplm has seven features above 50\%. Such a lower stability of steplm can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. For example, the first five selected models have a cumulative frequency of 76\% with projpred, but only of 14\% with steplm. In addition, we note that the size of the selected model is much smaller when using projpred.

@@

Table \ref{tab:model_performances} reports the predictive performances, in terms of root mean square error, of the full model and the corresponding selected submodel using both projpred and steplm. We also repeat the selection after adding truly uncorrelated (i.e., irrelevant) features up to a total of 100 covariates. For this experiment, we report both the size of the selected submodel and the number of selected irrelevant features. Whilst, with the original data, we do not observe any gain using projpred in terms of predictive performance, what we do observe is a smaller size for the selected submodel, at the cost of a higher computation time compared to steplm. However, when additional irrelevant features are added, projpred is able to notably improve the predictive performance keeping the same submodel size, whereas steplm selects a much larger submodel also including a large number of irrelevant features (13 irrelevant out of a total of 19 selected features).

@@

\hypertarget{computational-burden}{%
\subsection{Computational burden} \label{computational-burden}}

We consider the reference model approach as a family of methods, thus providing a general answer to the question of computational burden is not possible because not unique. It is possible to identify a common and inevitable overhead, though, which is the cost of fitting the reference model. In the simple ``bare-bones'' reference model approach we have described above, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance. However, including all the available features can often be quite computationally demanding. In such a case, a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases, this does not always result in an appreciable speed-up. For example, when there are several predictive but mutually uncorrelated features, all of them  have to be included in the reference model anyway. Other alternatives could be fixing some parameters to their maximum marginal likelihood value \cite[typically referred as empirical Bayes;][]{casella1985introduction} or use a completely frequentist approach, if predictions are good enough.

The comparison example between projpred and steplm shows that such additional computational burden can be worth it. We found that the reference model approach (projpred) selected much sparser submodels, which can result in saving time and general costs using the selected model for future predictions. Table \ref{tab:model_performances} displays that there could even be an improvement in the predictive ability compared to the full model, thus justifying more additional computational costs in the selection procedure. More benefits obtainable through a reference model approach are provided in the experiments discussed in the upcoming section.

@@

In this section, we demonstrate the benefits of reference model approaches using the normal means problem as a benchmark. The normal means problem consists of estimating the (usually sparse) vector of means of a vector of normally distributed observations. Let us denote the dimensionality of the vector of means by $p$ and let $\{z_{j}\}_{j=1}^{p}$ be the vector of observations of the random variables $\{Z_{j}\}_{j=1}^{p}$. In the normal means problem, the task is to estimate the latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model:
\
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}) \quad j=1,..,p
\end{equation}
Note that it is equivalent to a linear regression where the design matrix is the identity matrix with the number of features being equal to the number of observations. This formulation can be found in practice at various places, 
for example, in the analysis of microarray data, where a large set of genes are tested in two groups of patients labeled as positive or negative to some disease \citep{paper:efron,efron2012large}. The objective of the analysis is to select the subset of genes statistically relevant to the disease. One common way to proceed is to compute the two-samples $t$-statistic for every gene separately.
After normalising these statistics, they then become the data, $Z$, in the normal means problem \eqref{eq:normal_means_problem}. For further details see the examples in \cite{paper:efron, efron2012large}.

>>>>

\begin{document}
\maketitle
\begin{abstract}
Feature selection, or more generally, model reduction is an important aspect of statistical workflow aiming to provide insights from data. While a vast number of feature selection procedures are available, it often remains unclear what are the underlying principles explaining why certain methods are working better than others. We investigate and discuss one of these principles in detail, which we identify with the use of a reference model. We demonstrate how reference models act as a noise-filters on the target variable by approximating its data generation mechanism.
To this end, we devise a simple reference model approach that can be used on top of any feature selection method. In several numerical experiments, we show how this translates into better and more stable feature selection independently of the specific selection procedure. 
These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.

@@

In statistical applications, one of the main steps in the modelling
workflow is covariate or feature selection, which is a special case of
model reduction. Variable selection may have multiple goals. 
First, if the covariates themselves are of interest, we can use
variable selection to infer which variables contain predictive
information about the target variable.
Second, as simpler models come with the advantages of reduced
measurement costs and improved interpretability, we may be interested in
finding the minimal subset of covariates which
still provides good predictive performance (or good balance between simplicity and
predictive performance). 
When the predictive capability is guiding the selection, the true
data generation mechanism of future data can be approximated either by
using the observed data directly or alternatively by using predictions
from a \emph{reference model} \citep{vehtari2012survey}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward 
regression \citep{venables2013modern,harrell2015regression}, the observed empirical data distribution is utilised as a
proxy of future data usually in combination with cross-validation
or information criteria to provide estimates for out-of-sample
predictive performance.
In contrast, reference model based methods approximate the future data generation
mechanism using the predictive distribution of a reference model,
which can be, for example, a full-encompassing model including all
covariates. The reference
model approach has been used for long time in Bayesian statistics
since the seminal work of \citet{paper:reference_lindley}. For more
historical references, see papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for
most recent methodological development see 
\citet{paper:projpred}.

Reference models have been also used in non-Bayesian contexts,
where \cite{harrell2015regression} describes them as full models that
can be thought of as a ``gold standard'' (for a given application).
For example,
\cite{faraggi2001understanding} deal with the necessity of identifying
interpretable risk groups in the context of survival data using neural
networks, which typically perform very well in terms of prediction,
but whose covariates are difficult to be understood in terms of
relevance. 
\cite{paul2008preconditioning}, using the term preconditioning,
explore approximating models fitting Lasso or stepwise regression
against consistent estimates $\hat{y}$ of a reference model instead of
the observed responses $y$.

In this paper, we examine the benefits of using reference
models in detail. We illustrate how a properly designed reference model is able 
to filter parts of the noise present in the data, and hence provide an improved and more
stable selection process. This holds even if the reference model does not perfectly resemble the
true data generating process. Moreover, our analyses indicate that the substantial reduction of
variance attributable to noise is usually
more important than small potential bias due to model misspecification.
We argue that, regardless of how the reference model is set up and used in the
inference procedure, it can be always seen as acting as a filter on the observed 
data. Following this idea, we devise a simple reference model approach that can
be applied on top of any variable selection procedure.

In our simulations and analysis of real data, we apply several state-of-the-art methods for variable selection and compare,
for each of these methods, the results obtained by using a data based and a
reference model based approach. When a full 
variable selection is carried out, we report and compare quantities such as 
the false discovery rate \citep{paper:fdr_BH}, sensitivity and stability of the selection \citep{paper:stability}. For fully Bayesian methods, we additionally report results in terms of shrinkage and accuracy of the posterior estimates.
We show the benefits
of a reference model approach, regardless of what specific procedure is applied, using 
the normal means estimation problem as a benchmark. 
Variable selection is typically most challenging in 
\textit{small n large p} scenarios, that is, scenarios with small number of
observations but a lot of unknown model parameters. 
These occur frequently, for example, 
in microarray data analysis in medicine or biology \citep{dudoit2003multiple,liao2007logistic,paper:efron}.
In general, the less data we have and the more complex 
the estimated models, the higher the benefit of using a reference
models as the basis for variable selection.
Our results indicate that the core reason why the reference model based
methods perform well is the reference model itself, rather than the
specific way of using it. We do recognise that all these methods
diversify how the subset of variables is actually chosen and which
properties it has, thus it remains fundamental to investigate and
develop such different procedures. Although we focus on
Bayesian approach in this paper, the benefits of the reference model 
approach can also be obtained in non-Bayesian settings.

@@

The main assumption under any reference model approach is that we
operate in an $\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is we have constructed a
model which reflects in the best way our beliefs about the future
data. In practice, this would often be the full-encompassing model
(including all models we consider having non-zero prior
probability as submodels) which has passed model checking and criticism \citep[see,
e.g.][]{gelman2013bayesian}.

A good predictive model is able to filter part of the noise present in the data. The noise is the main source of the instability and tends to obscure the relevance of the covariates in relation to the target variable of interest. Indeed, let us consider the following explanatory example taken from \cite{paper:projpred}. The data generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim N(0,1) && \nonumber \\ 
     Y|&f\sim N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} N(\sqrt{\rho}f,1-\rho) \quad &&j=1,..,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} N(0,1) &&j=k+1,..,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy observation. The first $k$ covariates are strongly related to the target variable $Y$ and correlated among themselves. Precisely, $\rho$ is the correlation among any pair of the first $k$ covariates, whereas $\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between any relevant covariate and, respectively, $f$ and $Y$. If we had an infinite amount of observations, the sample correlation would be equal to the true correlation between $X_j$ and $Y$. However, even in this ideal asymptotic regime, this  correlation would still remain a biased indicator of the true relevance of each covariate (represented by the correlation between $X_j$ and $f$) due to the intrinsic noisy nature of $Y$.

@@

% In Figure \ref{fig:correlation} the reference model filters the noise.
% This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
% Above the correlation was used to estimate the relevance of a covariate.  
As an another example, let us consider a candidate submodel in a feature selection process and let us refer to the parameter distribution of the submodel by $\pi$ and to the induced predictive distribution by $q_{\pi}(\tilde{y})$. We would like to choose $\pi$ so that $q_{\pi}(\tilde{y})$ maximises some predictive performance utility, for example, the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y},
\end{equation}
where $p_{t}(\tilde{y})$ denotes the (usually unknown) true generating mechanism of future data $\tilde{y}$. If we refer to the posterior predictive distribution of a reference model with $p(\tilde{y}|D)$, where $D$ stands for the data on which we conditioned on, we can approximate \eqref{eq:elpd} using $p(\tilde{y}|D)$ instead of the true data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the Kullback-Leibler (KL) divergence from the reference model's predictive distribution to the submodel's predictive distribution:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred to as the projection of the predictive distribution, which is the general idea behind the projection predictive approach \cite[see][]{paper:projpred}. Again the reference model is acting as a filter, or substitute, on data.

@@

Relying on the idea of the reference model as a noise-filter, a generic reference model approach, which can be used on top of any feature selection procedure, consists of simply substituting the target observations with corresponding point predictions of the reference model, for instance, the posterior predictive means. Afterwards, any method for feature selection can be used. We do not argue that such a reference model approach should necessarily be used as is for feature selection in real applications. There are several more sophisticated reference model approaches available in the literature \cite[see, e.g.][]{paper:projpred,paananen2017variable,paul2008preconditioning}, which are likely to yield even better results. Rather, the reason for focusing on this simple version is to enable a fair comparison between a reference model and a purely data based approach when otherwise using the same selection procedures. Other kinds of comparisons, as the one presented in Section \ref{projection}, work well as motivational examples but lack in terms of generality since the presence or absence of a reference model is not the only aspect differentiating the two procedures. We will come back to the more general comparison in Section \ref{comparison}.

@@

As a motivational example to the implications of using a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior draws to the submodels \citep{paper:original_proj}. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Suppose we have observed $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates for which we want to obtain a minimally relevant subset. Than, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of $S$ draws from the reference model posterior.
\item Rank the covariates according to their relevance using some heuristics and consider as candidate submodels only those which preserve this order,  starting from including only the highest ranked covariate (the submodels are then naturally identified by their model size). This is not strictly necessary but reduces the number of submodels to consider in the following steps.
\item For each submodel, project the reference model posterior draws as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL} \left( p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\;||\;q(\tilde{y}_{i}|\boldsymbol{\theta}^{s}) \right), \quad s=1,..,S \,,
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the predictive distribution of the reference model with parameters fixed at $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariate values related to the statistical unit (identified by the subscript $i$), whereas $q(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ is the predictive distribution of the submodel. The projected draws $\boldsymbol{\theta}_{\perp}^{s}$ then present the projected posterior for the submodel.
\item For each submodel (size), test the predictive performance, for example, via cross-validation.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model's predictive utility score.
\end{enumerate}

@@

\subsubsection{Example: Variable selection in bodyfat prediction}

We show the benefits of the reference model in the form of the projection approach (projpred) with respect to the stepwise backward regression (steplm) using the body fat data \citep{johnson1996fitting}\footnote{The analysis is based on \texttt{R}-notebook by Vehtari available at \url{https://avehtari.github.io/modelselection/bodyfat.html}.}. The target variable of interest is the amount of body fat, which is measured with a complex and expensive procedure. Additionally, we have information about 13 covariates which are anthropometric measurements (e.g., height and weight), some of which are highly correlated among themselves and thus provide some challenges in the model selection. In total, we have 251 observations. The goal is to find the simplest model which is able to predict the amount of fat sufficiently well. We use the same data again later on in Section \ref{real-world-data}.

\cite{paper:bodyfat} report results using steplm with a significance level of 0.157 with AIC selection \citep{akaike1974new}, fixing abdomen and height to be always included in the model. Here, we implement steplm via AIC without fixing any variable. For the selection via projpred, the reference model includes all the covariates, we regularise model complexity with a regularised horseshoe prior on the covariate coefficients \citep{paper:rhs}, submodels are explored using forward search, and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. The experiments are implemented in \texttt{R} \citep{Rcore2018}. The steplm approach is carried out combining the \texttt{step} and \texttt{lm} functions, whereas the selection via projpred is done with the \texttt{projpred} package. In order to automatise the latter, we select the submodel as the one with the smallest size which has an elpd score higher (i.e., better) than the reference model score with probability of at least 5\%. The inclusion frequencies of each feature in the final model after 100 bootstrap samples are depicted in Figure \ref{fig:inclusion_frequencies}. The projpred approach has only two variables, `abdomen' and `weight', with inclusion frequencies above 50\% (`abdomen' is the only one included always), the third most included is `wrist' at 44\%, and the fourth is `height' at 35\%, while steplm has seven features above 50\%. Such a higher variability and lower stability of steplm can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. For example, the first five selected models have a cumulative frequency of 76\% with projpred, but only of 14\% with steplm. In addition, the size of the selected model is much smaller when using projpred.

@@

First two rows of Table \ref{tab:model_performances} report the predictive performances, in terms of root mean square error (RMSE), of the full model and the selected models using either the projection predictive approach or the stepwise backward elimination.
There is no difference in predictive performance of the selected models by different approaches even if there is difference in the selected features.

We repeat the experiment with a modified data set by adding 84 unrelated noisy features. Last two rows of Table \ref{tab:model_performances} report RMSE, the size of the selected model and the number of selected noisy features using either the projection predictive approach or the stepwise backward elimination.
The projection predictive approach has similar predictive performance and the same submodel size as in the first experiment, whereas the stepwise regression has worse predictive performance and selects a much larger submodel including a large number of irrelevant features.

@@

\hypertarget{computational-burden}{%
\subsection{Computational cost} \label{computational-burden}}

There are various ways to implement a reference model approach. A common term in the computational cost is the inference for the reference model. In the simple ``bare-bones'' reference model approach we have described above, it is the only added cost. In general, a full Bayesian reference model is advisable in order to have the best predictive performance. However, including all the available features can often be quite computationally demanding. In such a case, a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases, this does not always result in an appreciable speed-up. For example, when there are several predictive but mutually uncorrelated features, all of them  have to be included in the reference model. Other alternatives could be using point estimates such maximum a posteriori estimatge if predictions from such reference model without full Bayesian inference are good enough.

The comparison example between projpred and steplm shows that such additional computational cost can be worth it. We found that the reference model approach (projpred) selected much sparser submodels, which can result in improved and more stable interpretability and saving time and general costs using the selected model for future predictions. Table \ref{tab:model_performances} and results by \citet{paper:model_selection} demonstrate also that in case of large number of features, there can be a significant improvement in the predictive ability compared to feature selection without a reference model, thus justifying additional computational costs in the selection procedure. More benefits obtainable through a reference model approach are provided in the experiments discussed in the upcoming section.

@@

In this section, we analyse in more detail the benefits of reference model approaches using the normal means problem as a benchmark. The normal means problem consists of estimating the (usually sparse) vector of means of a vector of normally distributed observations. The dimensionality of the vector of means is denoted by $p$ and $\{z_{j}\}_{j=1}^{p}$ is the vector of observations of the random variables $\{Z_{j}\}_{j=1}^{p}$. In the normal means problem, the task is to estimate the latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model:
\
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}), \quad j=1,..,p.
\end{equation}
This is equivalent to a linear regression where the design matrix is the identity matrix with the number of features being equal to the number of observations. This formulation can be found in practice, 
for example, in the analysis of microarray data, where a large set of genes are tested in two groups of patients labeled as positive or negative to some disease \citep{paper:efron,efron2012large}. The objective of the analysis is to select the subset of genes statistically relevant to the disease. One common way to proceed is to compute the two-sample $t$-statistic for every gene separately.
After normalising these statistics, they then become the data, $Z$, in the normal means problem \eqref{eq:normal_means_problem}. For further details see the examples in \cite{paper:efron, efron2012large}.

In our experiments, we retrieve the normal means problem from the sample correlations between the target and the covariates using the Fisher $z$-transformation \citep{hawkins1989using}. Suppose we have a continuous target random variable $Y$, a set of $p$ continuous covariates $\{X_{j}\}_{j=1}^{p}$, and denote $\rho_{j}=Cor(Y,X_{j})$. Further, suppose we have observed $n$ statistical units and define $r_{j}$ the sample correlation between the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the $j$-th covariate $\{x_{ij}\}_{i=1}^{n}$. Finally, we refer to the Fisher $z$-transformation function $\text{tanh}^{-1}(\cdot)$ as $T_{F}(\cdot)$. Assuming each pair $(Y,X_{j})$ to be bivariate normally distributed, the corresponding transformed correlations are approximately normally distributed with known variance:
\
\begin{equation} \label{eq:fisher_transformation}
T_{F}(r_{j})\overset{ind}{\sim} N(T_{F}(\rho_{j}),\frac{1}{n-3}) \quad j=1,..,p.
\end{equation}
Therefore, rescaling the quantities $T_{F}(r_{j})$ by $\sqrt{n-3}$ and denoting the results as $z_{j}$, we have the formulation \eqref{eq:normal_means_problem} of the normal means problem, this time with unit variance:
\
\begin{equation} \label{eq:normal_means_problem2}
Z_{j}|\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \quad j=1,..,p.
\end{equation}
Now the quantities of interest $\theta_{j}$ are equal to $\sqrt{n-3}\,T_{F}(\rho_{j})$. There are different ways to proceed with the selection in the normal means problem. Here, we consider several state-of-the-art approaches primarily from Bayesian statistics.
